---
title: "Setting Up R Assignment"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---
Analyze the role of descriptive statistics in data exploration phase of analytics projects.

1. Install the ISLR library using the install.packages() command. Call the library using the library(ISLR) command to ensure that the library is correctly installed.
```{r}
require("ISLR") #loads the required package
library("ISLR") #activates the library

#install.packages("ISLR") - Since the package is already installed calling it off so that there won't be any issue while running the command.
```

2. Create a new R-Notebook (.Rmd) file. In the first code chunk, call the ISLR library and then print the summary of the Carseats dataset. How many observations (rows) this dataset contains?
```{r}
library("ISLR")
#activates the ISLR Package
summary(Carseats)#prints out the summary of the Carseats dataset which is a part of the ISLR Package
nrow(Carseats)#gives us the count of the total rows presented in the Carseats dataset
```

3. Using the summary statistics shown above, what is maximum value of the advertising attribute?
```{r}
max(Carseats$Advertising)
#since we have used summary function above using the max() function exclusively to find the max value of advertising attribute
```

4. Calculate the IQR of the Price attribute.
```{r}
IQR(Carseats$Price)
#IQR refers to the inter-quartile range
```

5. Plot the Sales against Price. What do you see in there? Calculate the correlation of the two attributes. What does the sign of the correlation coefficient suggest?
```{r}
#Using qplot function to plot sales vs price, before that we need to install and activate ggplot package
#install.packages("ggplot2") - If the package isn't installed.
library("ggplot2")
#Running the qplot function
qplot(data=Carseats,x=Sales,y=Price, xlab = "Sales", ylab = "Price", main = "Sales vs Price")
#using cor.test() function to calculate the correlation coefficient between the two variables Sales and Price.
cor.test(x = Carseats$Sales,y = Carseats$Price,method="pearson")
```

Key Findings from the plot Sales vs Price:
1. x and y variables have a negative or a inverse relationship.
2. Strength between the variables - They don't appear to be that strong, if they would have been strong then the points would have been closer helping to form an imaginary line.
3. Relationship - In here it appears that x and y have a linear relationship.

We see that the correlation coefficient value is -0.4449507 which falls under the bucket of -1 indicating that there exists a strong negative correlation or inverse relationship: this means that every time x increases, y decreases and vice-versa.

We are following the "pearson method" which is most widely used. It measures a linear dependence between two variables (x and y). It’s also known as a parametric correlation test because it depends to the distribution of the data. Alternatively we can change the method to "kendall" and "spearman" as well. 