---
title: "Online Retail Analytics"
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
  pdf_document: default
---
*Setting default values to get a clean output*
```{r}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

***Loading all the required packages***
```{r}
library("VIM")
library("ISLR")
library("caret")
library("class")
library("e1071")
library("ggplot2")
library("corrplot")
library("dplyr")
```


***Setting working directory and loading data***
```{r}
setwd("/Users/sampathnikhilkumar/Desktop")
data.df <- read.csv("Online_Retail.csv")
```


***1.Show the breakdown of the number of transactions by countries***
```{r}
data_country <- as.data.frame(table(data.df$Country))

data_country$Percentage <- data_country$Freq/nrow(data.df) * 100

colnames(data_country) <- c("Country", "Count", "Percentage")

data_country[data_country$Percentage > 1,]
```
*Countries accounting for more than 1% of the total transactions are EIRE, France, Germany and United Kingdom.*
\vspace{1mm}\newline

***2. Adding new attribute "TransactionValue" which is the product of Quantity and UnitPrice***
```{r}
data.df$TransactionValue <- data.df$Quantity * data.df$UnitPrice
```
*By adding this new attribute we can now calculate the value of the transactions based on our requirement.*
\vspace{1mm}\newline

***3. Using the newly created variable, TransactionValue, showing the breakdown of transaction values by countries with total transaction exceeding 130,000 British Pound***
```{r}
data.df %>% select(TransactionValue,Country) %>% group_by(Country) %>% summarise(Total = sum(TransactionValue)) %>% filter(Total >= 130000) %>% arrange((desc(Total)))
```
*There are total 6 countries where the transaction value exceeds 130,000 British Pound and the highest among them is "United Kingdom".*
\vspace{1mm}\newline

***4. Converting Invoice Date into a POSIXlt object***
```{r}
Temp=strptime(data.df$InvoiceDate,format='%m/%d/%Y %H:%M',tz='GMT')
head(Temp)

#New_Invoice_Date
data.df$New_Invoice_Date <- as.Date(Temp)

data.df$New_Invoice_Date[20000]- data.df$New_Invoice_Date[10]

#Invoice_Day_Week
data.df$Invoice_Day_Week= weekdays(data.df$New_Invoice_Date)

#New_Invoice_Hour
data.df$New_Invoice_Hour = as.numeric(format(Temp, "%H"))

#New_Invoice_Month
data.df$New_Invoice_Month = as.numeric(format(Temp, "%m"))
```

***4(a). Percentage of transactions (by numbers) by days of the week***
```{r}
data.df %>% group_by(Invoice_Day_Week) %>% summarise(count=n()) %>% mutate(percentage=count/nrow(data.df)*100)
```

***4(b). Percentage of transactions (by transaction volume) by days of the week***
```{r}
data.df %>% group_by(Invoice_Day_Week) %>% summarise(Total = sum(TransactionValue)) %>% mutate(Percentage = Total/sum(Total)*100)
```

***4(c). Percentage of transactions (by transaction volume) by month of the year***
```{r}
data.df %>% group_by(New_Invoice_Month) %>% summarise(Total = sum(TransactionValue)) %>% mutate(Percentage = Total/sum(Total) * 100)
```

***4(d). The date with the highest number of transactions from Australia***
```{r}
data.df %>% filter(Country =="Australia") %>% group_by(New_Invoice_Date) %>% summarise(Total_Count = n()) %>% arrange((desc(Total_Count)))
```
*As we can see from above on 2011-06-15 Australia has recorded the highest number of transactions i.e. 139 Transactions.*
\vspace{1mm}\newline

***5. Plot the histogram of transaction values from Germany***
```{r}
Germany <- data.df %>% filter(Country == "Germany") %>% group_by(New_Invoice_Month) %>% summarise(Total=sum(TransactionValue)) 

hist(Germany$Total, main = "Transaction Values of Germany", xlab="Total per month", ylab="Frequencies")
```

***6(a). Customer who had highest number of transactions***
```{r}
data.df %>% group_by(CustomerID) %>% select(CustomerID) %>% filter(!is.na(CustomerID)) %>% summarise(n_count = n()) %>% arrange((desc(n_count)))
```
*The CustomerID 17841 had the highest number of transactions amongst others with a total of 7983 transactions.*
\vspace{1mm}\newline

***6(b). Most valuable customer with the highest total sum of transactions***
```{r}
data.df %>% group_by(CustomerID) %>% select(CustomerID, TransactionValue) %>% filter(!is.na(CustomerID)) %>% summarise(Spending_max = sum(TransactionValue)) %>% arrange((desc(Spending_max)))
```
*The CustomerID 14646 is the most valuable customer with the highest spending sum of 279,489.020 British Sterling Pound.*
\vspace{1mm}\newline

***7. Percentage of missing values for each variable in the dataset***
```{r}
colMeans(is.na(data.df)*100)
```
*We can observe that CustomerID is the only attribute with 24.9266% of NAs in the entire dataset.*
\vspace{1mm}\newline

***8. The number of transactions with missing CustomerID records by Countries***
```{r}
data.df %>% filter(is.na(CustomerID)) %>% group_by(Country) %>% count()
``` 
*There are in total 8 countries and 1 unspecified country in the entire dataset which has NA values in them amongst these United Kingdom is the country with highest NA records of 133,600 rows.*
\vspace{1mm}\newline

***9. On average, how often the costumers comeback to the website for their next shopping?***
```{r}
Diff_Days <- data.df %>% select(CustomerID,New_Invoice_Date) %>% group_by(CustomerID) %>% distinct(New_Invoice_Date) %>% arrange(desc(CustomerID)) %>%  mutate(Days_Between = New_Invoice_Date-lag(New_Invoice_Date)) %>% filter(!is.na(Days_Between)) 

Diff_Days

mean(Diff_Days$Days_Between)
```
*On an average approximately for every 38 days customers come back to the website for their next shopping.*
\vspace{1mm}\newline

***10. Return rate of goods purchased by the customers from France***
```{r}
France_Cancel <- data.df %>% filter(Country=="France",Quantity<0) %>% count()

France_Total <- data.df %>% filter(Country=="France") %>% count()

Return_Percentage_France <- France_Cancel/France_Total*100
Return_Percentage_France
```
*The return rate of customers who made purchases in France is 1.741264%.*
\vspace{1mm}\newline

***11. The product that has generated the highest revenue for the retailer***
```{r}
data.df %>% select(StockCode,TransactionValue) %>% group_by(StockCode) %>%  summarise(Total=sum(TransactionValue)) %>% arrange((desc(Total))) 
```
*The product with the StockCode as "DOT" is the one which has generated highest revenue to the retailer i.e. 206,245.48 British Sterling Pound.*
\vspace{1mm}\newline

***12. Unique Customers in the dataset***
```{r}
data.df %>% select(CustomerID) %>% unique() %>% count()
```
*In total there are 4,373 unique customers in the dataset.*
